---
title: "Kaplan-Meier estimator"
author: "Thomas Lumley"
date: "12 September 2017"
output: html_document
---

First, read in the data

```{r}
pbc<-read.csv("pbc.csv")
head(pbc)
summary(pbc)
```

Load the `survival` package into your R session

```{r}
library(survival)
```

Now, fit the Kaplan-Meier estimator.  We package the failure time ($T$) and event status ($\Delta$) together into one thing with the `Surv` function.  Here, the event is `status==2`: a death, not a transplant. 

```{r}
pbc_surv <- survfit(Surv(time,status==2)~1, data=pbc)
pbc_surv
```

And plot it

```{r}
plot(pbc_surv)
```


Now, try customising the axis labels and lines. Use the help pages: eg `?survfit`, `?plot.survfit`.  Look at the Kaplan-Meier estimator for subsets of the data, such as people with high bilirubin.  

```{r}
high_bili <- subset(pbc, bili>2)
plot(survfit(Surv(time,status==2)~1,data=high_bili))
```

```{r}
plot(survfit(Surv(time,status==2)~I(bili>2),data=pbc))
plot(survfit(Surv(time,status==2)~I(bili>2),data=pbc),conf.int=TRUE,
     col=c("goldenrod","purple"),xscale=365,yscale=100)
```

Now, do the same things for the data in the file `colon.csv`.  There are two records for each person: time to recurrence of cancer and time, so you will want to make separate data sets for the subset of the records with `etype==1` (recurrence) and the subset with `etype==2` (deaths) and plot both curves.

```{r}
colon <-read.csv("colon.csv")
deaths<-subset(colon, etype==2)
recur<-subset(colon,etype==1)
plot(survfit(Surv(time,status)~1,data=deaths))
lines(survfit(Surv(time,status)~1,data=recur),col="red")
```
Can you get both curves on the same graph? (see `?lines.survfit` for hints)


