---
title: "Comparing groups"
author: "Thomas Lumley"
date: "12 September 2017"
output: html_document
---

Using the time-to-death variables in the colon cancer data set from the previous exercise, we're going to compare the treatment groups.  The `survdiff` function does logrank tests.


First, we'll plot the three curves for the three treatments

```{r}
library(survival)
colon <-read.csv("colon.csv")
deaths <- subset(colon, etype==2)
plot(survfit(Surv(time,status)~rx, data=deaths))
```

It looks like `Lev` doesn't do anyting, but that `Lev+5FU` is effective. Let's compare it to the other two

```{r}
survdiff(Surv(time,status)~I(rx=="Lev+5FU"), data=deaths)
```

Now, try comparing `Lev` to `Obs`: you'll need to get a subset of the data that doesn't have the `Lev+5FU` observations.

```{r}
survdiff(Surv(time,status)~I(rx=="Lev"), data=subset(deaths, rx!="Lev+5FU"))
```


Next, we want to look at quantiles

```{r}
deaths_km <- survfit(Surv(time,status)~rx,data=deaths)
deaths_km
```

Why is the median survival undefined for the `Lev+5FU` group?

```{r}
plot(deaths_km, col=c("goldenrod","purple","forestgreen"),conf.int=TRUE)
abline(h=0.5,lwd=2)
```

```{r}
quantile(deaths_km, probs=c(.1,.25,.5))
```

Now, look at survival percentages

```{r}
summary(deaths_km,times=365*c(1,2,5))
```

Is the difference in 5-year survival statistically significant? (There isn't a nice function for this)

```{r}
diff <- 0.634-0.526
se <- sqrt(0.0277^2+0.0282^2)
diff/se
```

Now do a similar comparison for five-year restricted mean

```{r}
print(deaths_km,rmean=5*365)
```

```{r}
diff<-1450-1339
se<-sqrt(33.0^2+33.4^2)
diff/se
```

The file `fakelogrank1.csv` has artificial data where the hazard rate curves cross.  Look at the survival curves for the two groups (with the time axis trimmed to a maximum of 10 to make it easier to see)

```{r}
fake<-read.csv("fakelogrank1.csv")
plot(survfit(Surv(time,status)~group,data=fake),xlim=c(0,10),col=c("black","red"))
```

Compare the logrank test and the weighted logrank tests with `rho=1` and `rho=-1`

```{r}
survdiff(Surv(time,status)~group,data=fake,rho=1)
survdiff(Surv(time,status)~group,data=fake,rho=-1)
```

If you are ok with installing packages, install the `muhaz` package and draw the hazard rate curves
```{r}
library(muhaz)
with(fake, plot(muhaz(time,status,subset=group==0,max.time=10)))
with(fake, lines(muhaz(time,status,subset=group==1,max.time=10),col="red"))
```