---
title: "Recurrent/clustered events"
author: "Thomas Lumley"
date: "10 September 2017"
output: html_document
---

## Recurrent events

The file `cgd.csv` contains data on 128 people (mostly children) with chronic granulomatous disease, an immune disorder characterised by the lack of interferon-$\gamma$. They were randomised to injections of interferon-$\gamma$ or placebo, and the event of interest is serious infection.

```{r}
library(survival)
cgd <- read.csv("cgd.csv")
```

- Fit a Cox model to estimate the effect of treatment (`treat`) and whether it differs according to the genetic cause (`inherit`). Remember to include `+cluster(id)` in the model formula to get the correct standard errors. 
- Repeat just using the first record for each person (these have `enum==1`). Compare the results
- Is there evidence that the treatment effect is different after the first event?
- Repeat, using time since the last infection rather than time on study as the time scale.

```{r}
coxph(Surv(tstart,tstop, status)~treat+cluster(id),data=cgd)
coxph(Surv(tstart,tstop, status)~treat*inherit+cluster(id),data=cgd)
coxph(Surv(tstart,tstop, status)~treat*inherit, data=subset(cgd, enum==1))
coxph(Surv(tstart,tstop, status)~treat*(I(enum==1)+inherit), data=cgd)
```

```{r}
library(survey)
m<-coxph(Surv(tstart,tstop, status)~treat*inherit+cluster(id), data=cgd)
regTermTest(m, ~treat:inherit,method="LRT")
```

```{r}
coxph(Surv(tstart-tstart,tstop-tstart, status)~treat*inherit+cluster(id),data=cgd)
coxph(Surv(tstop, status)~treat*inherit, data=subset(cgd, enum==1))
```



## Diabetic retinopathy

The file `retinopathy.csv` contains data on the treatment of diabetic retinopathy in 197 people with two eyes each. One eye was randomly chosen for laser treatment; the other was left untreated.

- Fit a Cox model for the effect of treatment with the eyes as matched pairs
- Does the effect of treatment vary by the laser type (`laser`) or by the type of diabetes (`type`)? A simple interaction model gives a warning here (why?). Use, eg, `~trt+trt:laser` rather than `~trt*laser` to avoid the warning. 
- Now look at the models considering the eyes just as clusters rather than matched sets. 

```{r}
ret<-read.csv("retinopathy.csv")
coxph(Surv(futime,status)~trt+trt:laser+strata(id),data=ret)
coxph(Surv(futime,status)~trt+trt:type+strata(id),data=ret)
```

```{r}
coxph(Surv(futime,status)~trt*laser+cluster(id),data=ret)
coxph(Surv(futime,status)~trt*type+cluster(id),data=ret)
```