---
title: "Competing risks"
author: "Thomas Lumley"
date: "10 September 2017"
output: html_document
---

The file `hypoxia.csv` has data on local and distance recurrence of cervical cancer in 109 women.  The main interest is in whether `ifp` and `hp5`, two measures of hypoxia (oxygen shortage) affect recurrence. We are interested only in the subset of women with initial complete response to treatment (`resp=="CR"`)

```{r}
hypoxia <- read.csv("hypoxia.csv")
```


- Fit a Cox model for time to recurrence of either sort (`dftime`, `dfcens`).
-  Fit Cox models for time to local recurrence, censoring distant recurrence (`dftime`, `failtype==1`) and time to distant recurrence censoring local recurrence (`dftime`, `failtype==2`), adjusting for `tumsize` and `pelvicln`. What do these mean?
- Why can't you meaningfully get survival curves from these models?

Now, install the `cmprsk` package and load it into your session.

First, look at the crude incidence curves for the two failure types overall, and by whether `ifp>20`
```{r}
library("cmprsk")
with(hypoxia, cuminc(dftime,failtype))
with(hypoxia, plot(cuminc(dftime,failtype)))
with(hypoxia, cuminc(dftime,failtype, ifp>20))
```

Now, we can do regression modelling
```{r}
with(hypoxia,crr(dftime,failtype,cbind(ifp,hp5,tumsize)))
```

Next, load the `pbc` dataset again and look at transplant (`status==1`) and death (`status==2`) as competing risks.